---
title: "USJudges_RMarkdown"
author: "fabianbohnenberger"
date: "25 September 2015"
output: html_document
---


## Aim

The aim of this document is to develop a weighted index of Judge's performance based on R's **"USJudgeRatings"** database. The dataset contains lawyers' ratings of State Judges in the US Superior Court. Additional information on the dataset can be found here: <https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/USJudgeRatings.html>

Data on four variables will be used: 

|variable| abbreviation|
|------|---|
|*Judicial integrity* | INTG|
|*Preparation for trial* | PREP|
|*Sound oral rulings* | ORAL|
|*Physical ability* | PHYS|


##Setting working directory 
```{r}
setwd("C:/Users/Fabian/Documents/GitHub/repo1/analysis/USJudges")

list.files()
```


##Descriptive Statistics 

### Overview

```{r, echo=FALSE}
## This section is calculating the descriptive values for the table below. 

#RANGE
rangeINTG <- range(USJudgeRatings$INTG)
rangePREP <- range(USJudgeRatings$PREP)
rangeORAL <- range(USJudgeRatings$ORAL)
rangePHYS <- range(USJudgeRatings$PHYS)

#MEDIAN
medianINTG <- median(USJudgeRatings$INTG)
medianPREP <- median(USJudgeRatings$PREP)
medianORAL <- median(USJudgeRatings$ORAL)
medianPHYS <- median(USJudgeRatings$PHYS)

## table is containing calculations for mean and sd (!) 
```

|variable|abbreviation|range|mean|median|standard deviation|
|--------|------------|-----|----|------|------------------|
|*Judicial integrity* |INTG |`r rangeINTG` | `r mean(USJudgeRatings$INTG)` | `r medianINTG` |`r sd(USJudgeRatings$INTG)`|
|*Preparation for trial* | PREP| `r rangePREP` |`r mean(USJudgeRatings$PREP)`| `r medianPREP` |`r sd(USJudgeRatings$PREP)`|
|*Sound oral rulings* | ORAL| `r rangeORAL` |`r mean(USJudgeRatings$ORAL)`| `r medianORAL` |`r sd(USJudgeRatings$ORAL)`|
|*Physical ability* | PHYS|`r rangePHYS` | `r mean(USJudgeRatings$PHYS)`| `r medianPHYS` |`r sd(USJudgeRatings$PHYS)`|


### Histograms 
```{r, echo=FALSE}
source('R_files/USJudges_Descriptive_Histograms.R')
## Histograms are sourced from another file
```



##Creating a weighted index

```{r, echo=FALSE}
#assigning observations to objects ($ = "component selector") 
x1 <- USJudgeRatings$INTG
x2 <- USJudgeRatings$PREP
x3 <- USJudgeRatings$ORAL
x4 <- USJudgeRatings$PHYS
```

###Defining weights for the Index
```{r}
w1 <- 0.25
w2 <- 0.25
w3 <- 0.25
w4 <- 0.25
```

###Creating function for weighted Index
- r identifies the row/Judge in the dataframe
- x1 = INTG, x2 = PREP, x3 = ORAL, x4 = PHYS
- w1 = weight for x1, ...
- Index will be adjusted to display an error message for numbers for r other than integers between 1 and 43 (see r source code)
```{r}
weighted_index <- function(r){
  x1[r]*w1+x2[r]*w2+x3[r]*w3+x4[r]*w4
}
``` 

```{r, echo=FALSE}
#adjusting Index to deal with numbers for r other than integers between 1 and 43
weighted_index <- function(r){
  if(1 <= r && r <= 43 && r == as.integer(r)){
    x1[r]*w1+x2[r]*w2+x3[r]*w3+x4[r]*w4
  } 
  else{
    print("Judge does not exist")
  }
}
```


###Results  
```{r}



#weighted_index(1:43) #for all Judges

#The index rating for Judge `r row.names(USJudgeRatings[0,1]` is `r weighted_index(1)`.
#The average rating for Judge `r row.names(USJudgeRatings[1,]` is `r weighted_index(1)`.
```


usjudges <- USJudgeRatings

row.names(usjudges)

row.names(usjudges])


`r row.names(USJudgeRatings[0,1]`


---

